---
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(tidy.opts = list(width.cutoff = 80), tidy = TRUE) # needs formatR package installed
```

## Overview

This document provides guidance on installing the dependencies for as well as running the experiments that can replicate the results in **Scalable Sampling of Truncated Multivariate Normals Using Sequential Nearest-Neighbor Approximation**.

## Installing Dependencies

Most of our dependencies are on CRAN, with only one exception. Please change `eval` to `TRUE` if you need to install the dependencies.
```{r, eval=FALSE, message=FALSE, results='hide'}
CRAN_pkg_names <- c("ggplot2", "fields", "scoringRules", "VeccTMVN", "sf", "spData", "GpGp", "nntmvn", "doParallel", "TruncatedNormal", "RColorBrewer", "autoimage", "mvtnorm", "RANN", "scales", "R.utils", "tidyr", "devtools")
Git_pkg_name <- "CensSpBayes"
Git_pkg_link <- "SumanM47/CensSpBayes"
for (pkg_name in CRAN_pkg_names) {
  install.packages(pkg_name, repos='http://cran.us.r-project.org')
}
```

The package `CensSpBayes` is not on CRAN. To install `CensSpBayes`, you need to first install [`INLA`](https://www.r-inla.org/download-install). The procedure for installing `INLA` depends on your operating system. There might also be several dependencies for `INLA` that cannot be installed through R.
```{r, eval=FALSE, message=FALSE, results='hide'}
library(devtools)
Git_pkg_name <- "CensSpBayes"
Git_pkg_link <- "SumanM47/CensSpBayes"
devtools::install_github(Git_pkg_link)
```

In fact, I could not install INLA on my MacOS, for which the `create_inla_mats` function cannot be called. The experiments involving `CensSpBayes` were run on a Linux system.

## Low Dimensional Experiments in Section 3

To replicate the results shown in Figure 1 and Table 1, run `lowdim_SNN.R`, `lowdim_SNN.R`, and `lowdim_SNN.R`.

To replicate the time growth curve, please run `time_compare.R` and `time_SNN.R`. Note that the reported times may be different from those shown in Figure 2.

## Diagnostic Plot for CSB

The diagnostic plot for CSB, shown in Figure D.1 is produced by running `CSB_diagnosis.R`. Note that the successful installation of `CensSpBayes` and `INLA` is required.

## High Dimensional Experiments in Section 4

Figures 3 as well as E.2, E.3, E.4 are each obtained by running `highdim_truth_known.R` and `highdim_truth_unknown.R` for 60 times each. Specifically, 

- Figure 3 and E.2 need the results from 1) running `highdim_truth_known.R` with `use_snn_order` set to 2, `scene_ID` sequentially set to 1, 2, and 3, and `k` sequentially set to $1, \ldots, 20$, which amount to a total of 60 runs and 2) running `highdim_truth_unknown.R` with `reorder` set to 1, `scene_ID` sequentially set to 1, 2, and 3, and `k` sequentially set to $1, \ldots, 20$, which amount to a total of 60 runs 
- Figure E.3 and E.4 need the results from 1) running `highdim_truth_known.R` with `use_snn_order` set to 0, `scene_ID` sequentially set to 1, 2, and 3, and `k` sequentially set to $1, \ldots, 20$, which amount to a total of 60 runs and 2) running `highdim_truth_unknown.R` with `reorder` set to 0, `scene_ID` sequentially set to 1, 2, and 3, and `k` sequentially set to $1, \ldots, 20$, which amount to a total of 60 runs 

The outputs from these runs were consolidated into `mtd_cmp.csv`. 

Figure E.5 is based on the results by running `m_compare_truth_known.R` for 180 times, corresponding to `use_snn_order` set to 0, 1, and 2; `scene_ID` set to 1, 2, and 3; and `k` set to $1, \ldots, 20$. The results were consolidated into `m_cmp.csv`.

Notice that the data simulation was done only once and the simulated data is stored in the `data` folder.

Each run of `highdim_truth_known.R` and `highdim_truth_unknown.R` will produce one set of samples and the corresponding heatmaps. From the many runs of `highdim_truth_known.R` and `highdim_truth_unknown.R`, a randomly chosen set of heatmaps was used as Figure~4.

## PCE Data Example in Section 5

Figures 5 and 6 in Section 5 can be reproduced by running `PCE_modeling.R`, where the control variables, `run_parm_est`, `run_VeccTMVN`, `run_TN`, `run_CB`, `run_seq_Vecc`, `run_seq_Vecc_all`, and `run_CB_all` should be set to `TRUE`.
 
Certain components of this script can take relatively long time

- `run_parm_est` is the switch for covariance parameter estimation, which took around 10 hours. Its result was store in `data/PCE_modeling.RData`. You can reuse the previous covariance estimation by setting `run_parm_est` to `FALSE` 
- `run_CB` and `run_CB_all` are the switches for using the CSB method for the inference over Texas and the entire US, respectively, the latter of which may take 3 hours
- All other methods are relatively fast
- The plotting part of `PCE_modeling.R` can only run after the results from `run_VeccTMVN`, `run_TN`, `run_CB`, `run_seq_Vecc`, `run_seq_Vecc_all`, and `run_CB_all` are all available












